\section[DiffEqn]{Differential Equations}
\begin{frame}{Slope field}

\begin{example}
p584-6
\end{example}



\begin{example}
The stock price is modeled by the following differential equation
\[
\frac{\mathrm{d}P}{\mathrm{d}t}=-P^{2}+2P+\sqrt{P}
\]
At which price level the price increase fastest?
\end{example}

\end{frame}

\begin{frame}{Solving equations}

\begin{example}
(separate equations) p575-12, 30.
\end{example}

\end{frame}

\begin{frame}{Exponential model}

\begin{example}
The bacteria in a certain culture increase continuously at a rate
proportional to the number present. If the number triples in $6$
hours, how many will there be in $15$ hours?
\end{example}



\begin{example}
Radium-226 decays at a rate proportional to the quantity present.
Its half-life is 1612 years. How long will it take for one quarter
of a given quantity of radium-226 to decay?
\end{example}

\end{frame}

\begin{frame}{Newton's law of cooling}

\begin{example}
A hot object cools at a rate proportional to the difference between
its own temperature and that of its environment. If roast at room
temperature $68^{\circ}\mathrm{F}$ is put into a $20^{\circ}\mathrm{F}$
freezer, and if, after $2$ hours, the temperature of the roast is
$40^{\circ}\mathrm{F}$:
\begin{enumerate}
\item What is its temperature after $5$ hours?
\item How long will it take for the temperature of the roast to fall to
$21^{\circ}\mathrm{F}$?
\end{enumerate}

\end{example}

\end{frame}

\begin{frame}{Logistic growth}

\begin{example}
Because of limited food and space, a squirrel population cannot exceed
\alert<2>{$1000$}. When there were $100$ squirrels $2$ years ago,
the \alert<2>{increase rate} of the population is $10$ squirrels
per year.
\begin{enumerate}
\item Find the \alert<2>{relative growth rate} of this population and write
down the differential equation.
\item Sketch the solution curve of the initial condition given above.
\item Use \alert<2>{Euler's method} to estimate the population now by 2
steps.
\end{enumerate}

\end{example}

\end{frame}



